Missing The Point

### 2020-10-15 Manuscript.Rmd
Restructure results section

# sim results plotting function
generic_plot <- function(v, name, ...){
  results %>%
    mutate(
      p = factor(p, levels = c(0.25, 0.5, 0.75), labels = c("25%", "50%", "75%")),
      mech = factor(mech, levels = c("MCAR", "MAR", "MNAR"))
      ) %>% 
  ggplot(aes(
    x=.data[["it"]], 
    y=.data[[v]], 
    color = .data[["p"]], 
    shape = .data[["mech"]], 
    linetype = .data[["mech"]]
    )) +
  geom_line(na.rm = TRUE) +
  labs(
    x = "Number of iterations",
    y = name,
    color = "Incomplete \n cases", 
    shape = "Missingness \n mechanism", 
    linetype = "Missingness \n mechanism") +
  theme_classic() + 
  scale_color_manual(values = c(
    #"#009E73", #green
    "#56B4E9", #blue
    #"#E69F00", #orange
    "#F0E442", #yellow
    "#D55E00"  #red
  ))
  # scale_color_manual(values = c(
  #   # '#228833', #green
  #   # '#66CCEE', #blue
  #   '#CCBB44', #yellow
  #   '#EE6677',  #red~ish
  #   '#AA3377'  #maroon
  #   ))
}
```

@@

- What does non-convergence look like? --> non-stat + non-mix in a certain parameter, e.g., chain means.

- What are the consequences? --> bias, under-coverage --> Refer to van Buuren (2018) instead of reproducing this

- Existing non-convergence identifiers for MCMC algorithms are not all applicable, e.g. N_eff is estimated using varcov matrix of the set of samples, while this is not the case in it im per se.

@@

We investigate non-convergence of the MICE algorithm through model-based simulation in R (version 4.0.2; R Core Team 2020). The number of simulation repetitions is 1000, and the simulation set-up is summarized in the pseudo-code below

```
# simulation set-up (pseudo-code)
for each simulation repetition (1:1000)
  1. simulate complete data
  for each missingness condition (1:9)
    2. create missingness
    for each iteration (1:50)
     3. impute missingness
     4. perform analysis of scientific interest
     5. compute non-convergence diagnostics 
     6. pool results across imputations
     7. compute performance measures
 8. combine outcomes of all missingness conditions
9. aggregate outcomes of all simulation repetitions
```

### Aims
Our aim is to assess the impact of non-convergence on the validity of scientific estimates obtained using `mice`. To induce non-convergence we terminate the iterative algorithm at different imputation chain lengths.

### Data generating mechanism
Data are generated in each simulation repetition for a sample of $n_{obs} = 1000$ cases. Let the complete set be three multivariate normally distributed random variables

$$
\begin{pmatrix}
Y\\
X_1\\
X_2
\end{pmatrix}
\sim \mathcal{N}
\begin{bmatrix}
\begin{pmatrix}
0\\
0\\
0
\end{pmatrix},
\begin{pmatrix}
2     &       &  \\
0.5   & 1     &  \\
-0.5  & 0.5   & 1
\end{pmatrix}
\end{bmatrix}.
$$
The complete set is amputed according to a $3\times3$ factorial design, consisting of three missingness mechanisms and three proportions of incomplete cases. [Briefly describe MCAR, MAR, MNAR + 25, 50, 75% of cases incomplete (not using 5% anymore, because according to Vink, n.d., 25% is more representative of missing data problems in the social sciences)]

### Estimands
We perform multiple linear regression to predict outcome variable $Y$: $Y' = \beta_0 + \beta_1 X_1 + \beta_2 X_2$, where $Y'$ is the expected value of the outcome. Our estimands are regression coefficient $\beta_1$, and the coefficient of determination, $r^2$.

### Methods
We use a combination of two non-convergence identifiers---autocorrelation and $\widehat{R}$---and four parameters of interest---chain means, chain variances, a scientific estimate, and the novel parameter we propose, $\lambda$.

### Performance measures
As recommended by @buur18, our performance measures are bias, empirical coverage rate and confidence interval width of the estimands.

<!-- - outcomes: estimated regression coefficient, CIW, coverage, r2 multivariate is key, because univariate convergence does not guarantee multivariate convergence -->

<!-- - convergence parameters: chain means, chain variances, scientific parameter, lambda -->

<!-- - diagnostics: rhat, ac -->

## Some results

```{r echo=FALSE}
generic_plot(v = "est", name = "Regression estimate") + 
  list(geom_hline(yintercept = 1, color = "grey"))

@@

Bias in estimate depends mostly on the amount of missingness (less info to estimate the relation??), but overall MNAR performs the worst, which is stable after it=5.

Update with different CI computation, see https://stefvanbuuren.name/fimd/sec-evaluation.html#sec:quantifyingbias

```{r echo=FALSE}
generic_plot( v ="cov", name = "Coverage rate") + 
  list(geom_hline(yintercept = 0.95, color = "grey"))

# generic_plot( v ="cov", name = "Coverage rate") + 
#   list(geom_hline(yintercept = 0.95, color = "grey")) +
#   ylim(0.85, 1)

Cov is more or less fine, except for MNAR with only 0.05% incomplete cases --> due to bias in the estimate in combination with very narrow CI (see CIW!). However, all stable after it=3??

```{r echo=FALSE}
generic_plot(v="CIW", name = "CI width")

@@

```{r echo=FALSE}
generic_plot(v="rsq", name = "Coefficient of determination (r^2)") +
  list(geom_hline(yintercept = 0.5, color = "grey"))

@@

```{r echo=FALSE}
generic_plot(v="ac.max.mu.Y", name = "AC chain means")

This plot above and the next three plots show autocorrelations with different thetas

```{r echo=FALSE}
generic_plot(v="ac.max.sigma.Y", name = "AC chain variances")

```{r echo=FALSE}
generic_plot(v="ac.max.qhat", name = "AC scientific estimate")

```{r echo=FALSE}
generic_plot(v="ac.max.lambda", name = "AC novel theta")

```{r echo=FALSE}
generic_plot(v="r.hat.max.mu.Y", name = "R hat chain means")

This plot above and the next three plots show r hat values with different thetas

```{r echo=FALSE}
generic_plot(v="r.hat.max.sigma.Y", name = "R hat chain variances")

```{r echo=FALSE}
generic_plot(v="r.hat.max.qhat", name = "R hat scientific estimate")

```{r echo=FALSE}
generic_plot(v="r.hat.max.lambda", name = "R hat novel theta")

@@

@@

<!-- ## Check if heuristic for AC works -->
<!-- ```{r} -->
<!-- load("Data/diagnostics.Rdata") -->
<!-- load("Data/raw_outcomes.Rdata") -->
<!-- sim_nr <- 3 -->
<!-- outcomes[[sim_nr]] %>% filter(mech == "MCAR", p == 0.95) %>% ggplot(aes(x=it, y = est)) + geom_line() -->
<!-- convergence_diagnostics %>% filter(sim == sim_nr, mech == "MCAR", p ==0.95) %>% ggplot(aes(x=it, y = ac.max.qhat)) + geom_line() -->
<!-- convergence_diagnostics %>% filter(sim == sim_nr, mech == "MCAR", p ==0.95) %>% ggplot(aes(x=it, y = ac.max.lambda)) + geom_line() -->
<!-- ``` -->

methodological explanation is that rhat and ac have a lag (few it to inform your statistic) --> will always indicate conv slower than the est

>>>>

# sim results plotting function
generic_plot <- function(v, name, ...){
  results %>%
    mutate(
      p = factor(p, levels = c(0.25, 0.5, 0.75), labels = c("25%", "50%", "75%")),
      mech = factor(mech, levels = c("MCAR", "MAR", "MNAR")),
      est = est - 1,
      rsq = rsq - 0.5
      ) %>% 
  ggplot(aes(
    x=.data[["it"]], 
    y=.data[[v]], 
    color = .data[["p"]], 
    shape = .data[["mech"]], 
    linetype = .data[["mech"]]
    )) +
  geom_line(na.rm = TRUE) +
  labs(
    x = "Number of iterations",
    y = name,
    color = "Incomplete \n cases", 
    shape = "Missingness \n mechanism", 
    linetype = "Missingness \n mechanism") +
  theme_classic() + 
  scale_color_manual(values = c(
    #"#009E73", #green
    "#56B4E9", #blue
    #"#E69F00", #orange
    "#F0E442", #yellow
    "#D55E00"  #red
  ))
  # scale_color_manual(values = c(
  #   # '#228833', #green
  #   # '#66CCEE', #blue
  #   '#CCBB44', #yellow
  #   '#EE6677',  #red~ish
  #   '#AA3377'  #maroon
  #   ))
}
```

@@

-   What does non-convergence look like? --\> non-stat + non-mix in a certain parameter, e.g., chain means.

-   What are the consequences? --\> bias, under-coverage --\> Refer to van Buuren (2018) instead of reproducing this

-   Existing non-convergence identifiers for MCMC algorithms are not all applicable, e.g. N\_eff is estimated using varcov matrix of the set of samples, while this is not the case in it im per se.

@@

<!-- The aim of the simulation study is to determine the impact of non-convergence on the validity of statistical inferences, and to assess whether non-convergence may be detected using several diagnostic methods. Inferential validity is reached when estimates are both  unbiased and have nominal coverage across simulation repetitions (nsim = 1000). -->

We investigate non-convergence in iterative imputation through model-based simulation in R (version 4.0.2; R Core Team 2020). We provide a summary of the simulation set-up in Algorithm 1; the complete script and technical details are available from [github.com/hanneoberman/MissingThePoint](https://github.com/hanneoberman/MissingThePoint).

    # simulation set-up (pseudo-code)
    for each simulation repetition (1:1000)
      1. simulate complete data
      for each missingness condition (1:9)
        2. create missingness
        for each iteration (1:50)
         3. impute missingness
         4. perform analysis of scientific interest
         5. compute non-convergence diagnostics 
         6. pool results across imputations
         7. compute performance measures
     8. combine outcomes of all missingness conditions
    9. aggregate outcomes of all simulation repetitions

### Aims

Our aim is to assess the impact of non-convergence on the validity of scientific estimates obtained using `mice` [@mice]. Inferential validity is reached when estimates are both  unbiased and have nominal coverage across simulation repetitions ($n_{sim} = 1000$). To induce non-convergence we terminate the iterative algorithm at different imputation chain lengths ($n_{it} = 50$). We differentiate between nine different missingness scenarios: three missingness mechanisms versus three proportions of incomplete cases [remove here if already defined above/below??].

### Data generating mechanism

Data are generated in each simulation repetition for a sample of $n_{obs} = 1000$ cases. Let the complete set be three multivariately normal random variables

$$
\begin{pmatrix}
Y\\
X_1\\
X_2
\end{pmatrix}
\sim \mathcal{N}
\begin{bmatrix}
\begin{pmatrix}
0\\
0\\
0
\end{pmatrix},
\begin{pmatrix}
2     &       &  \\
0.5   & 1     &  \\
-0.5  & 0.5   & 1
\end{pmatrix}
\end{bmatrix}.
$$

The complete set is amputed according to a $3\times3$ factorial design, consisting of three missingness mechanisms and three proportions of incomplete cases. [Briefly describe MCAR, MAR, MNAR + 25, 50, 75% of cases incomplete (not using 5% anymore, because according to Vink, n.d., 25% is more representative of missing data problems in the social sciences)]

### Estimands

We impute the missing data five times ($m = 5$) using Bayesian linear regression imputation with `mice` [@mice]. On each imputed dataset, we perform multiple linear regression to predict outcome variable $Y$ from the other two variables

$$
Y' = \beta_0 + \beta_1 X_1 + \beta_2 X_2,
$$
where $Y'$ is the expected value of the outcome. Our estimands are the regression coefficient $\beta_1$ and coefficient of determination $r^2$ that we obtain after pooling across the imputations.

### Methods

We use a combination of two non-convergence identifiers---autocorrelation and $\widehat{R}$---and four parameters of interest---chain means, chain variances, a scientific estimate, and the novel parameter we propose, $\lambda$.

### Performance measures

As recommended by @buur18, our performance measures are bias ($E(\bar{Q})-Q$), average confidence interval width ($\bar{Q}_{LL} - \bar{Q}_{UL}$, where $LL$ and $UL$ denote the lower and upper limit of the 95% confidence interval respectively) and empirical coverage rate ($Pr(\bar{Q}_{LL} \geq Q \geq \bar{Q}_{UL})$) of the estimands.

## Results

We present the results of our simulation study visually because of the many simulation conditions. Within each plot, the results are split according to the missingness conditions (missingness mechanisms as line types, and proportion of incomplete cases as colors).

### Non-Convergence Identifiers

The next four plots show the autocorrelations in each parameter plotted against the number of iterations.

```{r echo=FALSE}
p <- generic_plot(v="ac.max.mu.Y", name = "AC chain means")
p %>% plotly::ggplotly()
```

This plot shows that irrespective of the missingness condition, autocorrelation in the chain means decreases rapidly until $it \geq 6$. This means that there is some initial trending within chains, but the stationarity does not improve substantively after the first few iterations.

```{r echo=FALSE}
p <- generic_plot(v="ac.max.sigma.Y", name = "AC chain variances")
p %>% plotly::ggplotly() 
```

Autocorrelations in the chain variances do not show a systematic effect of the missingness conditions either. The number of iterations that is required to reach non-improving autocorrelations is somewhat more ambiguous.

```{r echo=FALSE}
p <- generic_plot(v="ac.max.qhat", name = "AC scientific estimate")
p %>% plotly::ggplotly() 
```

```{r echo=FALSE}
p <- generic_plot(v="ac.max.lambda", name = "AC novel theta")
p %>% plotly::ggplotly() 
```

The next four plots show r hat values with different thetas

```{r echo=FALSE}
p <- generic_plot(v="r.hat.max.mu.Y", name = "R hat chain means")
p %>% plotly::ggplotly() 
```

```{r echo=FALSE}
p <- generic_plot(v="r.hat.max.sigma.Y", name = "R hat chain variances")
p %>% plotly::ggplotly() 
```

```{r echo=FALSE}
p <- generic_plot(v="r.hat.max.qhat", name = "R hat scientific estimate")
p %>% plotly::ggplotly() 
```

```{r echo=FALSE}
p <- generic_plot(v="r.hat.max.lambda", name = "R hat novel theta")
p %>% plotly::ggplotly() 
```

### Performance Measures

```{r echo=FALSE}
p <- generic_plot(v = "est", name = "Bias in regression estimate") + 
  list(geom_hline(yintercept = 0, color = "grey")) 
p %>% plotly::ggplotly()
```

Bias in estimate depends mostly on the amount of missingness (less info to estimate the relation??). Overall 75% incomplete cases performs the worst, but is stable for $t \geq 6$.

<!-- Update with different CI computation, see <https://stefvanbuuren.name/fimd/sec-evaluation.html#sec:quantifyingbias> -->

```{r echo=FALSE}
p <- generic_plot( v ="cov", name = "Coverage rate") + 
  list(geom_hline(yintercept = 0.95, color = "grey"))
p %>% plotly::ggplotly()

# p <- generic_plot( v ="cov", name = "Coverage rate") + 
#   list(geom_hline(yintercept = 0.95, color = "grey")) +
#   ylim(0.85, 1)

@@

Cov is more or less fine, except for MNAR with only 0.05% incomplete cases --\> due to bias in the estimate in combination with very narrow CI (see CIW!). However, all stable after it=3??

```{r echo=FALSE}
p <- generic_plot(v="CIW", name = "CI width") 
p %>% plotly::ggplotly()
```

@@

```{r echo=FALSE}
p <- generic_plot(v="rsq", name = "Bias in coefficient of determination") +
  list(geom_hline(yintercept = 0, color = "grey"))
p %>% plotly::ggplotly()
```

@@

@@

- need for multivar parameters!!

@@

<!-- ## Check if heuristic for AC works -->

<!-- ```{r} -->

<!-- load("Data/diagnostics.Rdata") -->

<!-- load("Data/raw_outcomes.Rdata") -->

<!-- sim_nr <- 3 -->

<!-- outcomes[[sim_nr]] %>% filter(mech == "MCAR", p == 0.95) %>% ggplot(aes(x=it, y = est)) + geom_line() -->

<!-- convergence_diagnostics %>% filter(sim == sim_nr, mech == "MCAR", p ==0.95) %>% ggplot(aes(x=it, y = ac.max.qhat)) + geom_line() -->

<!-- convergence_diagnostics %>% filter(sim == sim_nr, mech == "MCAR", p ==0.95) %>% ggplot(aes(x=it, y = ac.max.lambda)) + geom_line() -->

<!-- ``` -->

methodological explanation is that rhat and ac have a lag (few it to inform your statistic) --\> will always indicate conv slower than the est
